---
title: 'Endo Ageing: beReady patients (age group vs receptivity status)'
author: 'Darina Obukhova'
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor options:
   chunk_output_type: console
---

```{r setuo, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align = 'center',
                      message=FALSE, warning=FALSE)
options(width = 1200)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggsci)
library(polycor)
```

```{r}
beready_pat <- read.csv2("~/projects/Endometrial_Ageing/BReady_patients_all.csv", header=T)

#Bin into the age groups

beready_pat <- beready_pat %>% 
  mutate(
    age_group = case_when(
      Age <= 29 ~ "20s",
      Age > 29 & Age <= 39 ~ "30s",
      Age > 39 & Age <= 44 ~ "40s",
      Age >= 45 ~ "45+"
    ),
    age_group = factor(
      age_group,
      level = c("20s", "30s", "40s", "45+")
    )
  )

#Samples per group

beready_pat %>% count(age_group)

#Distribution of samples
freq_table <- table(beready_pat$age_group, beready_pat$Reported)
percentage_table <- prop.table(freq_table, margin = 1) * 100 
percentage_df <- as.data.frame(percentage_table)
colnames(percentage_df) <- c("age_group", "reported_status", "percentage")
```

```{r}
# Create the plot
dist <- ggplot(percentage_df, aes(x = age_group, y = percentage, fill = reported_status)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Age group", y = "Percentage (%)", fill = "Receptivity status") +
  scale_fill_aaas() +
  theme(aspect.ratio = 1,
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line.y = element_line(colour="black"),
        axis.line.x = element_line(colour="black"),
        axis.text.x = element_text(angle = 30, vjust = 0.5, size = 12),
        axis.text.y = element_text(size = 10),
        legend.position = "bottom", 
        legend.spacing.y = unit(0.5, "cm"))
dist

ggsave("~/projects/Endometrial_Ageing/figures/beReady_stacked.pdf", 
       device = "pdf", height = 10, width = 15, units = "cm")
```

```{r}
recep <- ggplot(percentage_df, aes(x = reported_status, y = percentage, fill = age_group)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkorchid4", "seagreen4", "blue4", "darkorange2")) +
  labs(x = "Receptivity status", y = "Percentage (%)", fill = "Age group") +
  theme(aspect.ratio = 0.5,
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line.y = element_line(colour="black"),
        axis.line.x = element_line(colour="black"),
        axis.text.x = element_text(angle = 30, vjust = 0.5, size = 12),
        axis.text.y = element_text(size = 10),
        legend.position = "bottom", 
        legend.spacing.y = unit(0.5, "cm"))

recep

ggsave("~/projects/Endometrial_Ageing/figures/beReady_stacked_recepstatus.pdf", 
       device = "pdf", height = 10, width = 15, units = "cm")
```

Correlation between age group and receptivity status. As both of those variables are ordinal variable (ordinal scale assigns objects a name and arranges them in the order of their magnitude), we can try out polychoric correlation. It assumes that each ordinal data point comes from a normal distribution, and tries to estimate the correlation coefficient on that assumption.

```{r}
polychor(beready_pat$age_group, beready_pat$Reported, std.err=TRUE)
```

```{r}
cor.test(as.numeric(unclass(beready_pat$age_group)),as.numeric(unclass(beready_pat$Reported)), method="kendall")
```

Alternative age groups

```{r}

#Bin into alternative age groups

beready_pat <- beready_pat %>% 
  mutate(
    age_group_alt = case_when(
      Age <= 28 ~ "21-28",
      Age >= 29 & Age <= 36 ~ "29-36",
      Age >= 37 & Age <= 44 ~ "37-44",
      Age >= 45 ~ "45+"
    ),
    age_group_alt = factor(
      age_group_alt,
      level = c("21-28", "29-36", "37-44", "45+")
    )
  )

#Samples per group

beready_pat %>% count(age_group_alt)

#Distribution of samples
freq_table <- table(beready_pat$age_group_alt, beready_pat$Reported)
percentage_table <- prop.table(freq_table, margin = 1) * 100 
percentage_df <- as.data.frame(percentage_table)
colnames(percentage_df) <- c("age_group", "reported_status", "percentage")
```

```{r}
dist_alt <- ggplot(percentage_df, aes(x = age_group, y = percentage, fill = reported_status)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Age group", y = "Percentage (%)", fill = "Receptivity status") +
  scale_fill_aaas() +
  theme(aspect.ratio = 1,
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line.y = element_line(colour="black"),
        axis.line.x = element_line(colour="black"),
        axis.text.x = element_text(angle = 30, vjust = 0.5, size = 12),
        axis.text.y = element_text(size = 10),
        legend.position = "bottom", 
        legend.spacing.y = unit(0.1, "cm")) 
dist_alt

ggsave("~/projects/Endometrial_Ageing/figures/beReady_stacked_altgroups.pdf", 
       device = "pdf", height = 10, width = 15, units = "cm")

```

```{r}

recep_alt  <- ggplot(percentage_df, aes(x = reported_status, y = percentage, fill = age_group)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("darkorchid4", "seagreen4", "blue4", "darkorange2")) +
  labs(x = "Receptivity status", y = "Percentage (%)", fill = "Age group") +
  theme(aspect.ratio = 0.5,
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line.y = element_line(colour="black"),
        axis.line.x = element_line(colour="black"),
        axis.text.x = element_text(angle = 30, vjust = 0.5, size = 12),
        axis.text.y = element_text(size = 10),
        legend.position = "bottom", 
        legend.spacing.y = unit(0.5, "cm"))

recep_alt

ggsave("~/projects/Endometrial_Ageing/figures/beReady_stacked_recepstatus_altgroups.pdf", device = "pdf", height = 10, width = 15, units = "cm")
  
```

```{r}
polychor(beready_pat$age_group_alt, beready_pat$Reported, std.err=TRUE)
```

```{r}
cor.test(as.numeric(unclass(beready_pat$age_group_alt)),as.numeric(unclass(beready_pat$Reported)), method="kendall")
```

Does occurrence of specific receptivity status increases as age increases?

###Pre-receptive
```{r}

beready_pat_pr <- beready_pat[beready_pat$Reported == "pre-receptive", ]
cases <- as.vector(table(beready_pat_pr$age_group, beready_pat_pr$Reported))
total <- as.vector(table(beready_pat$age_group))
names(cases) <- names(table(beready_pat$age_group))
names(total) <- names(table(beready_pat$age_group))
prop.test(cases, total)

prop.trend.test(cases, total)

```
###Early receptive

```{r}

beready_pat_er <- beready_pat[beready_pat$Reported == "early-receptive", ]
cases <- as.vector(table(beready_pat_er$age_group, beready_pat_er$Reported))
total <- as.vector(table(beready_pat$age_group))
names(cases) <- names(table(beready_pat$age_group))
names(total) <- names(table(beready_pat$age_group))
prop.test(cases, total)

prop.trend.test(cases, total)
```

###Receptive
```{r}
beready_pat_r <- beready_pat[beready_pat$Reported == "receptive", ]
cases <- as.vector(table(beready_pat_r$age_group, beready_pat_r$Reported))
total <- as.vector(table(beready_pat$age_group))
names(cases) <- names(table(beready_pat$age_group))
names(total) <- names(table(beready_pat$age_group))
prop.test(cases, total)

prop.trend.test(cases, total)
```

###Late receptive
```{r}
beready_pat_lr <- beready_pat[beready_pat$Reported == "late-receptive", ]
cases <- as.vector(table(beready_pat_lr$age_group, beready_pat_lr$Reported))
total <- as.vector(table(beready_pat$age_group))
names(cases) <- names(table(beready_pat$age_group))
names(total) <- names(table(beready_pat$age_group))
prop.test(cases, total)

prop.trend.test(cases, total)
```
###Post-receptive

```{r}
beready_pat_pr <- beready_pat[beready_pat$Reported == "post-receptive", ]
cases <- as.vector(table(beready_pat_pr$age_group, beready_pat_pr$Reported))
total <- as.vector(table(beready_pat$age_group))
names(cases) <- names(table(beready_pat$age_group))
names(total) <- names(table(beready_pat$age_group))
prop.test(cases, total)

prop.trend.test(cases, total)

```

```{r}

beready_pat_pr <- beready_pat[beready_pat$Reported == "receptive", ]
cases <- as.vector(table(beready_pat_pr$age_group_alt, beready_pat_pr$Reported))
total <- as.vector(table(beready_pat$age_group_alt))
names(cases) <- names(table(beready_pat$age_group_alt))
names(total) <- names(table(beready_pat$age_group_alt))
prop.test(cases, total)

prop.trend.test(cases, total)

```


```{r}
actub <- read.csv2("~/projects/Endometrial_Ageing/Ac-Tub-a_per_LE_area.csv", header=T)

actub_df <- data.frame(
  value=c(actub$YMA, actub$AMA),
  group=factor(rep(c("YMA", "AMA"), each = 8))
)

actub_df$group <- factor(actub_df$group, levels=c('YMA', 'AMA'))

actub_markers_plt <- ggplot(data = actub_df , aes(x=group, y = value, fill=group))+
  stat_boxplot(geom = 'errorbar', width = 0.6) +
  geom_boxplot(width = 0.6) +
  labs(x=NULL, y="AcTubA stained MCC cells per LE") +
  scale_fill_manual(values = c("darkorchid4", "darkorange2"))+
  geom_jitter(width=0.25, alpha = 0.2) +
  theme_bw() +
  theme(aspect.ratio = 1.5,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_blank())
  
actub_markers_plt 

ggsave("~/projects/Endometrial_Ageing/figures/actub_mcc_le.pdf", plot = last_plot(), dpi = 1000)
```

```{r}
actub_pvals <- apply(actub_df[1], 2, function(x) t.test(x[1:8], x[8:16], exact = FALSE, paired=FALSE, p.adjust = "bonferroni")$p.value) 
```
